Can We Crunch the Social Graph on the Cheap?

### 2011-01-25 doc/hotpar2011/softxmt-hotpar2011.tex
finish up carls eval comments

\begin{figure*}[ht]
%list walk ILP
\begin{minipage}[b]{0.3\linewidth}
\centering
\begin{alltt}
  while (count-- > 0) \{
    list1 = list1->next;
    list2 = list2->next;
    \ldots
    list\(n\) = list\(n\)->next;
  \}
  \center{{\bf (a)}}
\end{alltt}
%\caption{Pseudocode for pointer chasing without coroutines.}
\label{fig:pointernocoro}
\end{minipage}
%listwalk green threads
\begin{minipage}[b]{0.35\linewidth}
\centering
\begin{alltt}
  while (count-- > 0) \{
     prefetch(&(list1->next));
     switch();
     list1 = read(&(list1->next));
 \}
  \center{{\bf (b)}}
\end{alltt}
  %%n lists
  % while (count-- > 0) \{
  % prefetch(&(list1->next));
  %   prefetch(&(list2->next));
   %  \ldots
   %  prefetch(&(list\(n\)->next));
   %  switch();
   %  list1 = read(&(list1->next));
   %  list2 = read(&(list2->next));
   %  \ldots
   %  list\(n\) = read(&(list\(n\)->next));
%\caption{Pseudocode for pointer chasing using coroutines.}
\label{fig:pointercoro}
\end{minipage}
%cache pressure
\begin{minipage}[b]{0.32\linewidth}
\centering
\begin{alltt}
  while (count-- > 0) \{
    prefetch(&(list1->next));
    switch();
    list1 = read(&(list1->next));
    for( i in 1 to num_local_updates ) \{
      local->data++;
      local = local->next;
   \}
  \}
   \center{{\bf (c)}}
\end{alltt}
%\caption{Pseudocode for pointer chasing with local updates.}
\label{fig:pointerupdate}
\end{minipage}
\caption{Pseudocode for: (a) pointer chasing without coroutines, (b) pointer chasing using coroutines, (c) pointer chasing with local updates.}
\label{fig:code}
\end{figure*}

@@

To find the maximum reference rate obtainable using our coroutine
library, we modified our pointer chasing benchmark as shown in
figure~\ref{fig:code}b. Now there are two sources of offered
concurrent references: the ILP exploited by the processor, and the
memory concurrency enabled by prefetching and switching to a new
coroutine. As with our first experiment, we allocated the lists in the
same processor as the cores doing the traversal.

@@

To simulate the performance of pointer chasing in a
multi-node system, we modified the scheduler of our coroutine library
to include a delay before a coroutine can be rescheduled, imitating
the network transit delay even though we are referencing local
memory. Recently published measurements of Infiniband latencies
\todo{cite} suggest we might expect 1.1 $\mu$s source-to-destination
latencies between two computers connected back-to-back; we assume a
switch transit delay of 400 ns \todo{how many hops} in each direction and thus estimate a 3
$\mu$s round-trip delay. In this experiment, once a coroutine has
yielded, we use the core's timestamp counter to ensure it does not get
rescheduled for 3 $\mu$s.

@@

We ran these experiments with two sources of memory concurrency: each
core ran up to 167 coroutines, and each coroutine made between 1 and 4
concurrent references. The results are shown in
Figure~\ref{fig:network-delay}. We are still able to reach a
near-maximum rate of 273 million references per second, but at least 2
concurrent references per coroutine are required. Running with a single
reference per coroutine achieves 236 million references per
second. Any network interface would be connected through a QPI link,
which we measured to have a bandwidth limit of 175 million references
per second, so our runtime appears to have the capacity to chase pointers
in a remote memory.

>>>>

\begin{figure*}[ht]
%list walk ILP
\begin{minipage}[b]{0.3\linewidth}
\centering
\begin{alltt}
  while (count-- > 0) \{
    list1 = list1->next;
    list2 = list2->next;
    \ldots
    list\(n\) = list\(n\)->next;
  \}
  \center{{\bf (a)}}
\end{alltt}
%\caption{Pseudocode for pointer chasing without coroutines.}
\label{fig:pointernocoro}
\end{minipage}
%listwalk green threads
\begin{minipage}[b]{0.35\linewidth}
\centering
\begin{alltt}
  while (count-- > 0) \{
     prefetch(&(list1->next));
     switch();
     list1 = read(&(list1->next));
 \}
  \center{{\bf (b)}}
\end{alltt}
  %%n lists
  % while (count-- > 0) \{
  % prefetch(&(list1->next));
  %   prefetch(&(list2->next));
   %  \ldots
   %  prefetch(&(list\(n\)->next));
   %  switch();
   %  list1 = read(&(list1->next));
   %  list2 = read(&(list2->next));
   %  \ldots
   %  list\(n\) = read(&(list\(n\)->next));
%\caption{Pseudocode for pointer chasing using coroutines.}
\label{fig:pointercoro}
\end{minipage}
%cache pressure
\begin{minipage}[b]{0.32\linewidth}
\centering
\begin{alltt}
  while (count-- > 0) \{
    prefetch(&(list1->next));
    switch();
    list1 = read(&(list1->next));
    for( i in 1 to num_local_updates ) \{
      local->data++;
      local = local->next;
   \}
  \}
   \center{{\bf (c)}}
\end{alltt}
%\caption{Pseudocode for pointer chasing with local updates.}
\label{fig:pointerupdate}
\end{minipage}
\caption{Pseudocode for: (a) pointer chasing without coroutines, (b) pointer chasing using coroutines, (c) pointer chasing with local updates. Coroutine code, as in (b), might also include ILP like what (a) depends on for memory concurrency, e.g., by doing multiple prefetches to independent lists, switch, then multiple reads.}
\label{fig:code}
\end{figure*}

@@

To find the maximum reference rate obtainable using our coroutine
library, we modified our pointer chasing benchmark as shown in
Figure~\ref{fig:code}b. Recall that the baseline code in Figure~\ref{fig:code}a relied solely on ILP to achieve memory concurrency, which may not be abundant in real applications. Using coroutines, there are two sources of offered
concurrent references: the ILP exploited by the processor, and the
memory concurrency enabled by prefetching and switching to a new
coroutine. As with our first experiment, we allocated the lists in the
same processor as the cores doing the traversal.

@@

To simulate the performance of pointer chasing in a
multi-node system, we modified the scheduler of our coroutine library
to include a delay before a coroutine can be rescheduled, imitating
the network transit delay even though we are referencing local
memory. Recently published measurements of Infiniband latencies
\todo{cite} suggest we might expect 1.1 $\mu$s source-to-destination
latencies between two computers connected back-to-back; we assume a
switch transit delay of 400 ns \todo{how many hops? 1or3?} in each direction and thus estimate a 3
$\mu$s round-trip delay. In this experiment, once a coroutine has
yielded, we use the core's timestamp counter to ensure it does not get
rescheduled for 3 $\mu$s.

@@

We ran these experiments with two sources of memory concurrency: each
core ran up to 167 coroutines, and each coroutine made between 1 and 4
concurrent references. The results are shown in
Figure~\ref{fig:network-delay}. We are still able to reach a
near-maximum rate of 273 million references per second, but at least 2
concurrent references per coroutine are required. Running with a single
reference per coroutine achieves 236 million references per
second. Note that the throughput keeps increasing up to around 850 total concurrent references offered, far more than the 36 we know the chip to support. This is because we have modeled the network latency by forcing coroutines to wait the extra time before using the data, so the physical miss buffers on our machine are not tied up any longer than usual.

We also observe that to achieve a throughput equal to the expected max outgoing network bandwidth (100 million references per second), \todo{reference earlier section calculation?} when there is only one reference per context available, we require 320 coroutines. This is close to the minimum of 300 concurrent references needed to tolerate $3 \mu s$ latency.

Any network interface would be connected through a QPI link,
which we measured to have a bandwidth limit of 175 million references
per second, so our runtime appears to have the capacity to chase pointers
in a remote memory.
